% Abstract: keeps the registered abstract's structure, framing phrases, and
% closing sentence (problem -> mechanism -> when-it-transfers -> profile ->
% prediction -> protection -> framework). The channel-conditional result is
% presented as the completed answer to the registered abstract's
% characterization sentence ("We further characterize when this local signal
% transfers across unlearning methods"), not as a new research question.
\begin{abstract}
An LLM asked to forget specific information should not lose unrelated
knowledge or capabilities that users still rely on. Yet approximate
unlearning can cause such collateral damage, while aggregate retention
metrics may obscure it when failures concentrate in a small, coherent
subset of retained behaviors. We identify update-conditioned gradient
interference as a mechanism underlying this failure: a retained behavior
is damaged in proportion to the coupling between its local sensitivity
and the realized parameter update. We further characterize when each
pre-unlearning signal transfers across unlearning methods, and find the
answer is structural: transfer is conditional on the \emph{damage
channel} induced by the unlearning objective, declared in advance by
where its loss reads the model. Objectives whose loss reads model outputs
(GA, GradDiff, NPO, SimNPO, and preference variants) damage candidates
with large loss-gradient magnitude, whereas objectives whose loss reads
internal representations (RMU, RepNoise, Circuit Breakers) damage
candidates whose hidden states lie close to the forget set; no single
score predicts well on both channels.

Guided by this mechanism, we construct candidate-level susceptibility
profiles without candidate-side backpropagation. For the output channel,
randomized symmetric perturbations -- two batched forward sweeps per
shared random direction -- estimate every candidate's gradient magnitude
at once and recover most of the exact gradient norm's predictive power;
for the representation channel, forward-only hidden-state proximity
suffices. Across balanced objective rosters, the channel-matched profile
reliably predicts per-example and tail damage along independently
configured trajectories, where channel-mismatched alternatives do not.
We then use the matched profile to allocate a fixed protection budget to
high-risk retained behaviors. At matched forgetting, channel-matched
protection reduces mean and worst-tail collateral damage on held-out
native audits, with comparable overall utility, relative to mismatched,
random, and no-protection allocation -- in a preregistered crossed design
over both channels. Together, these results support a probe-to-protection
framework that shifts LLM unlearning from post-hoc damage assessment
toward mechanism-informed pre-unlearning risk prediction and targeted
preservation.
\end{abstract}
