\section{Introduction}
\label{sec:intro}

A deployed language model may be asked to remove a person's private records,
a copyrighted document, or a harmful capability long after training has
finished. Retraining from scratch for every such request is usually
impractical, so machine unlearning suppresses the designated influence
through targeted post-training updates while preserving the model's
remaining behavior \cite{cao2015towards,bourtoule2021machine,
liu2024rethinking}. The selective requirement is what makes unlearning
hard: forget enough to satisfy the request, but not so aggressively that
useful knowledge is destroyed with it.

The resulting collateral damage is easy to miss because it concentrates.
A model can preserve most retained examples while failing sharply on a
small, coherent subset -- other authors when one author is deleted, benign
science that shares terminology with a suppressed hazardous procedure.
Split-aware evaluations expose this by reporting forget-adjacent retention
separately \cite{cheng2026entanglement,chang2025retain}, but they are
retrospective: the vulnerable behavior is identified only after unlearning
has been run and paid for. Practitioners would rather know \emph{before}
unlearning where damage will land, so that a limited protection or repair
budget can be spent where it is needed.

Prior attempts at such prospective prediction share a hidden assumption:
that susceptibility is a property of the model and data alone, so that some
single score -- lexical or embedding similarity to the forget set, gradient
alignment, gradient magnitude -- should rank vulnerable candidates
regardless of which unlearning method is applied. Empirically these scores
behave inconsistently across methods, and each new proposal appears to win
on some objectives and lose on others. This paper explains that
inconsistency and turns it into structure:

\begin{quote}
\emph{Retain-side susceptibility is not a model-intrinsic scalar. It is
conditional on the damage channel induced by the unlearning objective --
and the channel is declared in advance by where the objective's loss reads
the model.}
\end{quote}

To first order, the damage to a retained candidate $x$ under a parameter
update $\Delta\theta$ is $d(x)\approx g_x^{\top}\Delta\theta$: a coupling
between the candidate's sensitivity $g_x$ and the update. What
$\Delta\theta$ looks like is determined by the objective's loss. Objectives
whose loss is a function of model \emph{outputs} -- token likelihoods or
preferences, as in GA, GradDiff, NPO, SimNPO, IdkDPO, and GRU
\cite{yao2024large,zhang2024negative,fan2024simplicity,maini2024tofu,
wang2025gru} -- push probability mass, and the candidates that move with
them are those with large loss gradients. Objectives whose loss is a
function of internal \emph{representations} -- hidden states, as in RMU,
RepNoise, and Circuit Breakers
\cite{li2024wmdp,rosati2024repnoise,zou2024circuit} -- displace activation
geometry, and the candidates that move are those whose representations lie
close to the forget set, however small their loss gradients. The two
channels call for different predictor families, and we find the
correspondence empirically sharp in one direction: gradient-magnitude
probes light up on output-channel objectives and go dark on RMU, while
representation-proximity probes peak exactly there. No probe wins
everywhere; the winner flips with the channel.

Acting on this structure is only useful if the matching signal is cheap at
candidate-pool scale. Exact per-candidate gradient norms require one
backward pass per candidate. We instead estimate the entire pool's gradient
magnitudes with a \emph{backward-free randomized-sensitivity probe}:
symmetric forward evaluations at $\theta_0\pm\eta v_r$ along a few shared
random directions $v_r$ turn squared central differences into an unbiased
estimate of every candidate's squared gradient norm simultaneously. A
fidelity gate separates the estimator's numerical validity from its
predictive value, and the probe recovers most of the exact norm's
predictive power at a fraction of its cost. The representation channel is
cheaper still: forward-only hidden-state proximity to the forget set.

Prediction becomes protection through routing. Given a deletion request and
a declared objective, the channel determines the probe; the probe ranks the
discovery-fold candidates; the top-ranked pool receives a fixed repair
budget under a guarded procedure that caps forgetting regression. Our
success criterion is deliberately not a leaderboard: it is the
\emph{interaction} in a preregistered crossed design. For an output-channel
parent (GradDiff) and a representation-channel parent (RMU), we cross each
with matched, mismatched, random, and no-protection selectors and test
whether the channel-matched selector yields the lowest audit-tail damage at
matched forgetting -- per parent. If matching were unnecessary, the
mismatched and random arms would do as well; they do not.

Our contributions are threefold:
\begin{itemize}
    \item \textbf{Channel-conditioned susceptibility.} We show that which
    pre-unlearning signal predicts collateral damage depends on the
    unlearning objective's damage channel, declared in advance from its
    loss definition (outputs vs.\ representations). The evidence is an
    objective$\times$predictor-family interaction across balanced rosters
    of both channels, not a per-method anecdote.

    \item \textbf{A backward-free probe for the gradient channel.} A
    randomized symmetric-perturbation estimator profiles gradient
    magnitude over the whole candidate pool using only $2R$ batched
    forward sweeps, with a preregistered fidelity gate that certifies when
    the estimate is trustworthy under a given radius and precision.

    \item \textbf{Channel-matched protection.} Routing a fixed protection
    budget through the channel-matched probe reduces mean and worst-tail
    audit damage at matched forgetting, while mismatched and random
    routing do not -- establishing that the channel prediction is
    actionable, for any parent unlearner, at near-zero forgetting cost.
\end{itemize}

We position this work as a prediction-and-protection \emph{layer} on top of
existing unlearners, not as a competing unlearning objective. The remainder
of the paper follows this structure. Section~\ref{sec:related} organizes
prior unlearning methods by channel and identifies the channel-matching
gap. Section~\ref{sec:setup} formalizes damage channels, the prediction
and protection endpoints, and the preregistered interaction criterion
(Figure~\ref{fig:channel-mechanism}). Section~\ref{sec:method} presents
the probes, the fidelity gate, and the guarded channel-matched protection
procedure. Section~\ref{sec:experiments} reports the channel matrix, the
crossed protection experiment, cost, and cross-dataset replication.
