\section{Related Work}
\label{sec:related}

\paragraph{Evaluation granularity and localized damage.}
LLM-unlearning benchmarks already evaluate multiple facets of the
forget--retain trade-off, including utility, privacy leakage, and metric
robustness
\cite{maini2024tofu,shi2024muse,dorna2025openunlearning}. Yet methods and
checkpoints are commonly compared through set-level aggregates, and metric
aggregation or model selection can change which method appears strongest
\cite{thaker2024position,dorna2025openunlearning}. Fine-grained studies show
that damage concentrates in syntactically, semantically, or
representationally adjacent retain regions and can include benign failures
missed by static benchmarks
\cite{chang2025retain,cheng2026entanglement,ko2025holes}. These findings
establish that collateral damage is structured, but mainly localize it after
unlearning or within predefined neighborhoods. We instead rank a declared
universe of retain candidates before the target trajectory.

\paragraph{Objective families and damage geometry.}
Approximate objectives place their primary suppression pressure on output
likelihoods and preferences
\cite{jang2023knowledge,yao2024large,zhang2024negative,
fan2024simplicity} or directly on internal representations, as in RMU
\cite{li2024wmdp}. Related representation-control interventions have also
been adapted as unlearning baselines
\cite{rosati2024repnoise,zou2024circuit}. This readout taxonomy
does not partition causal mechanisms: output objectives also move hidden
states, and representation objectives ultimately change outputs.
Update-conditioned analyses further show that interference varies across
objectives, examples, steps, and layers \cite{wang2025gradient}. We therefore
use the readout only as a soft prior over complementary parameter-sensitivity
and representation-proximity signals.

\paragraph{Prospective risk.}
Adjacent methods estimate forget-side difficulty or set-level interactions
\cite{niu2025guard,chen2026hamu,cheng2026cud}, while trajectory analyses
quantify update-conditioned interference \cite{wang2025gradient}. They do
not resolve future risk for individual retain candidates. Recent work moves
closer to this target. One approach learns supervised
regressors from labeled outcomes of executed runs to predict a normalized
damage ratio for forget--evaluation-set pairs using cross-set features; its
auditor does not take target-model gradients as inputs
\cite{su2026preunlearn}. Concurrent work ranks non-target concepts using
semantic, parameter-footprint, gradient, and Jacobian measures, primarily
against NLL damage caused by gradient ascent; its prompt-disjoint and
behavioral probes show weaker transfer
\cite{anonymous2026neighbors}. Another proactive study predicts
algorithm-dependent corruption from model weights and activations
\cite{su2026corruption}.

Our distinction is therefore not the existence of pre-unlearning ranking.
Loss-shake energy measures a candidate's own direction-agnostic local loss
response rather than alignment with a selected forget update or overlap
between forget and candidate sensitivity supports. Unlike a footprint
score, it does not require candidate-side Fisher or gradient computation.
Its component score is computed directly at the target checkpoint without
fitting a supervised auditor, combined with forget-conditioned hidden-state
proximity across objective families, and evaluated at per-behavior
resolution. For protection, any deployable mixture weight is frozen on
separate development requests before the scored trajectory.

\paragraph{Mitigation and protection.}
Retention-aware methods mitigate damage by reweighting forget examples,
using similarity-informed constraints on the forget--retain trade-off,
curating forget-set coresets, or regularizing selected retain neighborhoods
\cite{niu2025guard,chen2026hamu,patil2025upcore,chang2025retain,
cheng2026entanglement}. These interventions change the unlearning data or
optimization. Our protection test instead holds the deletion request and
parent unlearner fixed, then uses the prospective profile only to allocate a
fixed repair budget. This separates risk prediction from the effect of
redesigning the unlearning objective.

\paragraph{Randomized sensitivity estimation.}
Influence and scalable data-attribution methods relate training examples to
model behavior through gradients or curvature
\cite{koh2017influence,park2023trak,kwon2024datainf}; forward-only variants
reduce candidate-side differentiation but estimate training-data value, not
future unlearning damage \cite{ko2024mirrored,deng2026forvalue}.
Symmetric perturbations and squared random projections are likewise
classical tools for recovering local gradient information from forward
evaluations \cite{spall1992spsa,malladi2023mezo,
hutchinson1989stochastic}. Our contribution is not this numerical machinery,
but the shared pool-wide forward scorer, its composition with
forget-conditioned proximity, and its use in a fixed-budget protection test.
