% 4.method.tex
% When Does LLM Unlearning Fail?
% Channel-conditioned susceptibility: probes, routing, guarded protection.

\section{Method}
\label{sec:method}

Section~\ref{sec:setup-channels} established which candidate property
couples to each damage channel. This section supplies the operational
pieces: a backward-free estimator of gradient magnitude for the output
channel (Section~\ref{sec:method-gradient}), a forward-only proximity
score for the representation channel
(Section~\ref{sec:method-representation}), the routing and pool
construction that turn a declared objective into a protect pool
(Section~\ref{sec:method-routing}), and the guarded repair procedure that
spends the protection budget without undoing forgetting
(Section~\ref{sec:method-protection}).

\subsection{Gradient-Channel Probe: Randomized Sensitivity}
\label{sec:method-gradient}

Let $B$ be a declared parameter block and
$g_x=\nabla_B\ell(x;\theta_0)$ the candidate gradient at the
pre-unlearning model. The output-channel target quantity is the
direction-agnostic sensitivity
\begin{equation}
    m(x)=\lVert g_x\rVert_2.
    \label{eq:exact-susceptibility}
\end{equation}
Computing $m(x)$ exactly requires one backward evaluation per candidate;
we retain this exact profile as the expensive reference. The scalable
estimator reuses symmetric forward perturbations. Let
$v_1,\ldots,v_R$ be seeded random unit vectors in $B$, and let
\begin{equation}
    \delta_\eta(x;v)
    =
    \frac{\ell(x;\theta_0+\eta v)-\ell(x;\theta_0-\eta v)}{2\eta}
    \label{eq:directional-probe}
\end{equation}
be the symmetric response of candidate $x$ along $v$. The randomized
sensitivity score is
\begin{equation}
    r_{\mathrm{rand}}(x)
    =
    \frac{1}{R}
    \sum_{r=1}^{R}
    \delta_\eta(x;v_r)^2 .
    \label{eq:randomized-susceptibility}
\end{equation}
Because $\mathbb E_v[(g_x^\top v)^2]=\lVert g_x\rVert_2^2/d_B$ for random
unit directions in a $d_B$-dimensional block, and the central difference
cancels the leading curvature term,
$r_{\mathrm{rand}}(x)\approx\lVert g_x\rVert_2^2/d_B$ up to Monte Carlo
and $O(\eta^2)$ error; the shared factor $d_B$ does not affect ranking.
The relative Monte Carlo variance is $2/R$ independently of $d_B$, so a
few dozen directions suffice. The directions are shared across the entire
candidate universe: $R$ directions cost $2R$ batched candidate-pool
forward sweeps in total, versus one backward evaluation \emph{per
candidate} for the exact profile. Memory never exceeds inference: no
per-example gradient is materialized.

\paragraph{Fidelity gate.}
The estimator is trusted only where a preregistered gate certifies it. The
radius $\eta$ must sit between two failure regimes: too small, and the
paired losses agree to within floating-point resolution so their
difference is cancellation noise; too large, and higher-order terms
contaminate the response. The gate first verifies the perturbation is
realized at all (measured $\lVert\eta v\rVert/\eta$ and the fraction of
changed block parameters, which catches low-precision underflow), then
decomposes agreement between three quantities on a diagnostic subset: the
exact squared norm $A=\lVert g_x\rVert^2$, the exact projected proxy
$B=(d_B/R)\sum_r (g_x^\top v_r)^2$, and the finite-difference estimate
$C$. Rank agreement of $B$ with $A$ isolates Monte Carlo error; rank
agreement of $C$ with $B$ isolates finite-difference error, including
cancellation. The radius and $R$ are frozen once the gate passes and
before any target trajectory is unsealed.

\paragraph{Rejected diagnostic: forget-direction alignment.}
The same probe evaluated along the normalized initial forget gradient
$\hat g_f$ yields the classical alignment score
$\delta_\eta(x;\hat g_f)\approx\langle g_x,\hat g_f\rangle$. We verified
this estimate against exact Jacobian--vector products, confirmed its
numerical validity, and preregistered its rejection as a susceptibility
predictor: across engines it correlates weakly or negatively with realized
damage, because a multi-step trajectory does not preserve its initial
direction. It appears in our tables as a cautionary baseline; the
supporting analysis is in the appendix.

\subsection{Representation-Channel Probe: Hidden-State Proximity}
\label{sec:method-representation}

The representation channel damages candidates whose hidden states occupy
the displaced region, so its matched probe is proximity, not gradient
magnitude. Candidate and forget examples are embedded with mean-pooled
final hidden states of the pre-unlearning model; each candidate is scored
by its mean cosine similarity to its $k$ nearest forget examples. The
score is forward-only, needs no parameter perturbation, and -- decisively
for the crossed design -- selects an almost disjoint candidate set from
the gradient probe: the two families' rankings are nearly uncorrelated on
our benchmarks. External sentence embeddings and lexical overlap provide
weaker members of the same family and are reported as comparators.

\subsection{Channel Routing and Profile Construction}
\label{sec:method-routing}

Given a deletion request and a declared parent objective, the router reads
the objective's channel from its loss definition
(Section~\ref{sec:setup-channels}) and selects the matched probe family:
randomized sensitivity for output-channel parents, hidden-state proximity
for representation-channel parents. Nothing in the router depends on
outcomes; it is a lookup on the preregistered channel table.

The selected scorer assigns one value to every candidate in the declared
universe. For a fixed budget $K_p$, the protect pool is
$\mathcal P_{K_p}(r)=\operatorname{TopK}_{x\in\mathcal C_{\mathrm{disc}}}\,r(x)$,
with deterministic tie-breaking; a size-matched low-score pool
$\mathcal R_0(r)$ is sampled from a frozen near-zero quantile of the
discovery fold and serves as the repair-time reference stream. Pool size,
quantile, group-disjoint folds, seed, and fallback rules are fixed before
target outcomes are observed, and no audit candidate can enter either
pool.

\subsection{Soft Channel Routing: The Channel-Mixture Score}
\label{sec:method-mixture}

The discrete router commits fully to one probe family. The spectrum
clause of Section~\ref{sec:setup-channels}, however, predicts residual
off-channel signal -- output-channel objectives perturb representations
incidentally -- so we generalize routing to a one-parameter family. Let
$\tilde r_g,\tilde r_h:\mathcal C\to[0,1]$ be the two headline probes
after rank normalization over the candidate universe (rank transforms
because the raw scales -- a squared-gradient estimate and a cosine -- are
not commensurable; the normalization is part of the frozen definition).
The \emph{channel-mixture score} is
\begin{equation}
    s_\alpha(x)
    =
    (1-\alpha)\,\tilde r_g(x)
    +
    \alpha\,\tilde r_h(x),
    \qquad \alpha\in[0,1],
    \label{eq:mixture}
\end{equation}
so that $\alpha$ is the weight on the representation route. The
endpoints recover the discrete router exactly -- $\alpha{=}0$ is the
randomized-sensitivity probe, $\alpha{=}1$ the proximity probe -- so
every result of the preceding sections is the $\alpha\in\{0,1\}$ special
case, and the declared channel now supplies a \emph{prior} over $\alpha$
rather than a hard switch: output-channel parents map to $\alpha{=}0$,
representation-channel parents to $\alpha{=}1$, and hybrid objectives
(e.g.\ RepNoise's auxiliary output term) to interior values.

Two disciplines keep the mixture honest. First, $\alpha$ is never chosen
on sealed data: it is either (i) the preregistered channel-to-$\alpha$
mapping, fixed before any outcome is read (primary -- the deployed
router is unchanged, and the mixture is preregistered as exploratory
instrumentation that does not reweight the primary endpoints), or
(ii) fit on discovery folds only (adaptive variant), with the sealed
audit untouched in both cases. Second, prediction and protection are allowed different
optima: the crossed experiment of Section~\ref{sec:exp-protection}
motivates measuring the prediction-optimal $\alpha^{*}_{\mathrm{pred}}$
(which maximizes damage correlation) separately from the
protection-optimal $\alpha^{*}_{\mathrm{prot}}$ (which minimizes sealed
audit tail damage when the pool is $\operatorname{TopK}$ by $s_\alpha$),
because repair generalization can propagate through representation-space
neighborhoods regardless of the parent's channel. The mixture sweep of
Section~\ref{sec:exp-mixture} reports both curves; how far the two
optima separate is itself an empirical finding, not an assumption.

\subsection{Guarded Channel-Matched Repair}
\label{sec:method-protection}

The budget is spent by an engine-then-repair pipeline that leaves the
parent unlearner untouched. The parent runs from $\theta_0$ until its
first saved checkpoint satisfying the common forget criterion; a parent
that never reaches remains a reach failure and is not repaired. At the
reaching state we cache per-sequence and per-token forget losses as paired
references, then minimize answer-token NLL on $\mathcal P_{K_p}$.

Repair must not silently undo forgetting. At fixed refresh intervals we
compute mean forget and reference-stream gradients and project the repair
update away from their span, making those mean losses locally invariant to
first order. Because a mean-level projection does not protect every forget
target, an identity-paired one-sided guard supplements it: with
$\bar u_i$ the cached post-unlearning forget loss and $u_i$ its current
value,
\begin{equation}
    D_{\downarrow}(u,\bar u)
    =
    \frac{1}{n}
    \sum_{i=1}^{n}
    \bigl[\bar u_i-\epsilon-u_i\bigr]_+^2
    \label{eq:guard}
\end{equation}
penalizes any per-example forgetting regression beyond tolerance
$\epsilon$, at sequence and (where enabled) token level. At audited
refreshes, a repair step that exceeds either paired-loss budget is
rejected, the last accepted parameters are restored, and the step size is
reduced. The output is the final saved repair checkpoint that still
satisfies the forget criterion; retention is evaluated only on the sealed
audit.

In the crossed design of Section~\ref{sec:setup-protection}, every
selector arm -- matched, mismatched, random -- runs this identical
pipeline with the identical budget; only the ranking that fills
$\mathcal P_{K_p}$ changes. Any difference between arms is therefore
attributable to where the budget went, which is exactly the channel
question.

\subsection{Scope of the Method}
\label{sec:method-scope}

The profile is a prospective risk measure, not a deletion certificate:
forgetting efficacy must still be established by direct, paraphrase, and
adversarial audits. The channel table covers where current objectives read
the model; an objective reading a third quantity would require a third
probe family, and hybrid objectives inherit partial signal from both. The
backward-free claim is scoped to candidate-side profiling: parents and
repair still backpropagate, as they must.
