% Evidence follows the three questions sealed in Section 3.

\section{Experiments}
\label{sec:experiments}

We evaluate the sealed framework defined in
Sections~\ref{sec:setup}--\ref{sec:method}, not a replacement unlearning
objective. Target profiles are fixed at $\theta_0$, parents run unchanged,
and only trajectories reaching the external forgetting condition enter the
two experiments. The evidence follows the research questions:
prospective damage ranking (RQ1), loss-shake fidelity and added value (RQ2),
and fixed-budget protection (RQ3). Figure~\ref{fig:channel-main} gives this
claim ladder; Tables~\ref{tab:core-evidence} and~\ref{tab:robustness} retain
the decisive effects, eligibility, and failure boundaries.

\input{figures/fig6_evidence_chain}

\subsection{Protocol and Comparison Contract}
\label{sec:benchmark-protocol}

\paragraph{Sealed splits.}
Candidates are split by semantic group into discovery and audit folds. Three
disjoint target-free development folds separately fix $(R,\eta)$, the
prediction weight, and the protection weight, together with $K_p$, repair
budget, stopping, eligibility, and tie rules. Target discovery candidates may
fit rank maps and supply repair examples; audit candidates never enter repair.
No target retained-damage or sealed-audit outcome changes optimization,
checkpoint selection, a weight, or a repair pool. The target direct-forgetting
metric alone selects $\theta_{t^\dagger}$, and final checkpoints use only the
broader forgetting/utility constraints. Guard and general-utility sets are
frozen, selector-independent, and disjoint from $\mathcal C_{\mathrm{audit}}$
and the native retained-behavior audit.

\paragraph{Settings and parents.}
The primary setting is TOFU single-author deletion
\cite{maini2024tofu}. Validation covers WMDP-bio/MMLU
\cite{li2024wmdp,hendrycks2021measuring}, MUSE-News
\cite{shi2024muse}, and RWKU \cite{jin2024rwku}; MUSE-Books and PISTOL
\cite{qiu2024pistol} are stress settings. Models include Qwen2.5-1.5B,
Qwen2.5-7B, and Llama-3.1-8B. The fixed parent roster spans GradDiff, NPO,
SimNPO, GRU, RMU, RepNoise, and Circuit Breakers, with GA and IdkDPO as stress
controls. Every dataset--model--request--parent cell and seed is declared in
advance; planned, completed, profile-valid, gate-reaching, feasible, and
claim-passing counts remain separate.

\paragraph{Predictor and allocation references.}
Prediction compares the frozen joint profile with $S_0$, $S_1$, exact gradient
energy, forget-update alignment, initial NLL, answer length, lexical and
sentence-embedding proximity, request-shuffled $\widehat q_H$, one-direction
loss response, and repeated frozen random rankings. The strongest simple
control is selected on development requests, never on the target. Protection
uses four confirmatory comparators---no repair, repeated-random allocation,
$S_0$, and $S_1$---under one common harness. Exact-energy allocation, the
prediction-weight selector, uniform rehearsal, and a realized-discovery-damage
oracle are diagnostics that separate estimator error, weight transfer, repair
capacity, and selector quality; they do not replace the confirmatory controls.

\paragraph{Measures.}
Damage is candidate-level $\Delta$NLL. We report request-level mean and
$\mathrm{CVaR}_{.95}$ plus each dataset's native retained-behavior metric. A
paired hierarchical bootstrap resamples target requests, seeds within
requests, and semantic groups. Correlations are formed within each
request--seed--parent cell; seeds are then averaged and requests receive equal
weight. Appendix~\ref{app:outcomes} gives denominator, CVaR, missingness, and
multiple-testing rules.

\subsection{RQ1: Prospective Damage Prediction}
\label{sec:exp-prediction}

For each eligible cell, let
$\rho_S=\operatorname{Spearman}(S_{\widehat\alpha_{\mathrm{pred}}},d)$ and
define paired endpoint gains
\begin{equation}
\begin{aligned}
  g_G&=\rho_S-\operatorname{Spearman}(S_0,d),\\
  g_H&=\rho_S-\operatorname{Spearman}(S_1,d),\\
  g_{\mathrm{ctl}}&=\rho_S-\operatorname{Spearman}(b^\star,d),
\end{aligned}
\label{eq:paired-prediction-gains}
\end{equation}
where $b^\star$ is the strongest predeclared simple control on the development
fold. The rank condition passes only when the one-sided 95\% lower bounds for
$\rho_S$, $g_G$, and $g_H$ are all positive; $g_{\mathrm{ctl}}$ separately
tests added value beyond simple controls. This absolute condition prevents a
joint score from passing merely by being less negative than both components.

Figure~\ref{fig:channel-main}A first tests whether aggregate retention hides a
concentrated, semantically coherent damage tail, then shows the prospective
effects. Let $m=K_p$ and $q=m/|\mathcal C_{\mathrm{audit}}|$.
$\mathrm{Recall}_q$ is the fraction of the $m$ largest positive-damage
candidates recovered by the profile's top $m$. Define the above-chance lift
$L_{\mathrm{tail}}=\mathrm{Recall}_q/q-1$. It is
secondary and is defined only when at least $m$ candidates have positive
damage. Such cells never leave the RQ1 correlation denominator; the table
reports tail-eligible $n/N$. RQ1 passes only when the rank condition holds,
the 95\% lower bound for $L_{\mathrm{tail}}$ is positive, and tail coverage is
at least 80\%.

\paragraph{Headline finding.}
{\setlength{\emergencystretch}{3em}
\TailHeadline\ \PredictionHeadline\par}

Table~\ref{tab:core-evidence} reports the absolute joint correlation, both
endpoint gains, tail lift and coverage, and RQ1 eligibility/pass by parent.
Appendix Table~\ref{tab:prospective-prediction} expands the simple controls,
horizons, and construction checks. A favorable point estimate without a valid
profile, reached gate, shared support, and all required positive bounds is not
a claim pass.

\subsection{RQ2: Loss-Shake Fidelity and Added Value}
\label{sec:exp-fidelity}

RQ2 separates computation from usefulness. First,
Figure~\ref{fig:channel-main}B compares $\widehat q_G$ with exact streaming
$q_G^\star$ on the same block, pool, batches, precision, and device. We report
rank correlation, Top-$K_p$ overlap, perturbation survival, direction-split
stability, synchronized time, and peak memory over $R$ and $\eta$. The exact
reference streams one candidate gradient at a time; where feasible, a batched
per-example-gradient implementation is included as a timing sensitivity.
Loss-shake uses $2R$ pool-forward sweeps and no candidate-side reverse mode.
This access pattern, rather than guaranteed speed, is its invariant advantage.

Let $f_\rho=\operatorname{Spearman}(\widehat q_G,q_G^\star)$ and let $f_K$ be
their Top-$K_p$ overlap fraction. The outcome-blind engineering floors are
$\tau_\rho=0.80$ and $\tau_K=0.70$. RQ2 is eligible only with valid
perturbations, an exact-energy reference, and common support for the frozen
simple control. It passes only when the 95\% lower bounds for
$f_\rho-\tau_\rho$, $f_K-\tau_K$, $g_H$, and $g_{\mathrm{ctl}}$ are all
positive.

The target-cell gains in Eq.~\eqref{eq:paired-prediction-gains} then ask
whether the estimated sensitivity is decision-relevant. In particular,
$g_H>0$ measures what loss-shake adds to proximity; replacing
$\widehat q_G$ by exact $q_G^\star$ inside the same frozen profile separates
finite-difference error from failure of the gradient-energy concept.
Fidelity alone does not show that either component improves prediction.

\paragraph{Headline finding.}
{\setlength{\emergencystretch}{3em}\FidelityHeadline\par}

Table~\ref{tab:core-evidence} reports both fidelity-floor margins, $g_H$,
$g_{\mathrm{ctl}}$, and RQ2 eligibility/pass; a high-fidelity estimator that
adds nothing to proximity or fails to beat the frozen simple control does not
pass.
Appendix Table~\ref{tab:bwfree} reports pool examples/tokens, hardware,
precision, warm-up, repeated synchronized time, memory, and target validity
coverage. Appendix~\ref{app:estimator-stability} varies direction bank,
radius, block, pool size, and precision and treats low-precision perturbation
collapse as invalid rather than as a favorable zero response.

\subsection{RQ3: Fixed-Budget Protection}
\label{sec:exp-protection}
\label{sec:exp-alpha}

All active repair arms branch from the same $\theta_{t^\dagger}$ and share the
operator, neutral stream, example order, optimizer seed, processed-token
budget, projection, guards, and audit schedule; only the $K_p$ selected
discovery examples change. The prediction and protection weights are fitted
on different development folds, so this is a separate decision-value test,
not an assumption that prediction correlation causes repair success.

For comparator $a$ and outcome
$k\in\{\mathrm{mean},\mathrm{CVaR}_{.95}\}$, let
$\Delta_{a,k}=Y_{\mathrm{profile},k}-Y_{a,k}$, where lower damage is better.
Let $M_a$ be the favorably oriented native retained-behavior metric and define
$h_a=M_{\mathrm{profile}}-M_a+\delta_{\mathrm{nat}}$, with the dataset-specific
non-inferiority margin frozen before target runs. Protection passes only if
all eight one-sided 95\% upper bounds for $\Delta_{a,k}$ are negative, all four
lower bounds for $h_a$ are positive, and every arm lies in the same
setting-specific feasibility region. The four comparators are no repair,
repeated random, $S_0$, and $S_1$. We call this comparison
\emph{constraint-matched}: final forgetting scores need not be identical, and
utility is a shared floor rather than an unreported pairwise equivalence.
RQ3 is eligible only when the unrepaired entry checkpoint and every repair arm
have common-support outcomes and satisfy this full region; other reached runs
remain in the declared denominator but cannot pass RQ3.

\paragraph{Headline finding.}
{\setlength{\emergencystretch}{3em}\ProtectionHeadline\par}

Figure~\ref{fig:channel-main}C and Table~\ref{tab:core-evidence} report
proposed and no-repair mean/CVaR damage, the decisive paired bound, native
retained behavior, constraint slack, and update diagnostics. Budget sweeps and
oracle or exact-energy diagnostics in Appendix Table~\ref{tab:crossed} show
whether a failure comes from insufficient repair capacity, the estimator, or
the allocation rule.

\subsection{Breadth and Failure Boundaries}
\label{sec:exp-supporting}

Table~\ref{tab:robustness} keeps every predeclared setting and parent in the
denominator. We do not pool parents into a universal effect. A setting
supports the full chain only when at least one output-readout and one
representation-readout primary parent are eligible and pass RQ1--RQ3
after the predeclared within-readout correction. The paper-level breadth claim
additionally requires support in primary TOFU, one non-primary model
scale/family, and at least two of WMDP-bio, MUSE-News, and RWKU; stress settings
cannot rescue a failure.

\paragraph{Headline finding.}
{\setlength{\emergencystretch}{3em}\TransferHeadline\par}

We explicitly report profile invalidity, parent non-reach, prediction worse
than an endpoint, diffuse damage where random allocation wins,
infeasible repair arms, few accepted updates, and paraphrase or extraction
regressions. These boundaries determine the licensed claim: the framework
provides prospective and decision-relevant information beyond its components
and allocation controls only in settings that pass the declared contract.
