% appendix.tex
% When Does LLM Unlearning Fail?
% Channel-Conditioned Prediction and Protection of Susceptible Retained Behavior
%
% Structure (mirrors the main text):
%   A  Reproducibility: data, objectives, contract, inference
%   B  Probe implementation and the fidelity gate
%   C  Protection implementation (pools, repair, guard)
%   D  Supplementary prediction results (incl. alignment post-mortem)
%   E  Supplementary protection results (incl. strong-method context)
%   F  Cost, replication, and stress details
%   G  Additional limitations and reporting scope

\appendix

% ============================================================ A. Reproducibility

\section{Reproducibility: Data, Objectives, and Evaluation Contract}
\label{app:evaluation-contract}

This section records the experimental context required to reproduce the
main results. It introduces no additional headline claim. Unless stated
otherwise, every choice is frozen on development requests before
test-request scores or outcomes are inspected, and the preregistered
constants (channel table, probe settings, endpoints, dataset roster) are
fixed before the confirmatory runs are unsealed.

\subsection{Datasets, Models, and Frozen Splits}
\label{app:datasets}

Table~\ref{tab:datasets} lists every setting with its role in the channel
claim. The primary mechanism setting uses single-author TOFU requests at
full fp32 precision, where the fidelity gate certifies the randomized
probe; the 7B setting provides profiling cost and scale transfer.
Replication settings rerun the reduced crossed roster on datasets whose
forget--retain proximity structure differs; stress settings target the
declared honesty clauses.

\begin{table*}[t]
\caption{Dataset and model matrix. Requests within one trained target are
nested and are not treated as independent model replicates. Discovery
candidates may guide protection; native audits are sealed outcomes.
Replication and stress tiers use the reduced crossed roster
(\{GradDiff, NPO\}$\times$\{RMU, RepNoise\}) with the full selector set.}
\label{tab:datasets}
\centering
\small
\setlength{\tabcolsep}{4.5pt}
\begin{tabular}{@{}lllll@{}}
\toprule
Setting & Target model & Units & Native audit & Role \\
\midrule
TOFU single-author
    & Qwen2.5-1.5B (fp32)
    & 20 requests/target
    & predeclared author relations
    & confirmatory (channel matrix + crossed) \\
TOFU scale/cost
    & Qwen2.5-7B (bf16)
    & 6 requests
    & predeclared author relations
    & profiling cost, scale transfer \\
WMDP-bio
    & Qwen2.5-1.5B (fp32)
    & bio forget corpus
    & frozen MMLU subjects
    & replication (representation habitat) \\
MUSE-News
    & Qwen2.5-1.5B (fp32)
    & 5 requests
    & held-out news
    & replication (output habitat) \\
RWKU
    & Qwen2.5-1.5B (fp32)
    & 10 requests
    & neighbor probes
    & replication (pretrained knowledge) \\
MUSE-Books
    & Qwen2.5-1.5B (fp32)
    & 3 requests
    & held-out chapters
    & stress (verbatim memorization) \\
PISTOL
    & Qwen2.5-1.5B (fp32)
    & 10 requests
    & held-out relations
    & stress (maximal entanglement) \\
\bottomrule
\end{tabular}
\end{table*}

For every request, the complete candidate universe $\mathcal C$, its
semantic-group discovery and audit folds, the native audit, and a
uniform-random audit are frozen before scores or unlearning outcomes are
computed. Groups are defined at the author, entity, article, or
dataset-specific semantic unit so that directly related examples do not
cross the discovery--audit boundary; the relation rule, group IDs,
candidate counts, and audit membership are persisted in the frozen data
manifest. Only $\mathcal C_{\mathrm{disc}}$ is eligible for score-guided
selection or repair. Scores on $\mathcal C_{\mathrm{audit}}$ are computed
at $\theta_0$, sealed, and unavailable to hyperparameter selection,
candidate eligibility, optimizer configuration, checkpoints, or stopping.
Ranking ties are broken by a frozen candidate identifier.

Exact scalar gradient norms are computed for the declared candidate pool.
Full candidate gradient vectors, needed only for the diagnostic
magnitude--cosine decomposition of Appendix~\ref{app:alignment-postmortem},
are computed on a uniform subset
$\mathcal C_{\mathrm{grad}}\subseteq\mathcal C_{\mathrm{disc}}$ of
$\min(128,|\mathcal C_{\mathrm{disc}}|)$ candidates, sampled before
outcomes are inspected; this subset can affect neither the deployed
profile nor protection.

\subsection{Objectives and Channel Assignments}
\label{app:objectives}

The preregistered channel table assigns each parent objective from its
loss definition: output channel -- GA, GradDiff, NPO+retain, SimNPO,
IdkDPO, GRU; representation channel -- RMU, RepNoise, Circuit Breakers.
Each parent's objective, trainable block, retain stream, learning rate,
update budget, and saved-checkpoint grid are fixed on development requests
without susceptibility scores. Representation parents receive longer step
budgets because their losses move hidden states rather than likelihoods
and reach the forget criterion more slowly; budgets are fixed per parent
before scoring. Parents that fail to reach the criterion within budget are
retained in the prediction matrix (their columns measure weak-signal
regimes) but flagged, and are excluded from criterion-conditional
protection contrasts. Our repair trajectory is excluded from confirmatory
prediction.

\subsection{Development and Resource Contract}
\label{app:contract}

Each method receives the same declared development budget, and the grid is
not retuned per test request. Model selection may use forget behavior and
discovery-fold retention but never sealed scores, native retention, or
target-trajectory damage. Table~\ref{tab:resource-parity} reports
optimization resources; profiling cost is reported separately in
Table~\ref{tab:full-cost}.

\begin{table*}[t]
\caption{Optimization information and resource contract. Selector arms of
one parent share the same parent optimizer, processed retain tokens,
updates, and repair budget. Profiling cost is excluded here
(Table~\ref{tab:full-cost}). Extra reference or data access is disclosed
rather than normalized away.}
\label{tab:resource-parity}
\centering
\scriptsize
\setlength{\tabcolsep}{2.8pt}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lcccccccc@{}}
\toprule
Method & Selector & Extra reference/data & Trainable block &
Retain tokens & Unlearn updates & Repair updates &
Optim.\ GPU-h & Optim.\ peak mem. \\
\midrule
Initial target reference
    & --- & full-data target & none & --- & --- & --- & --- & \tblph \\
Retain-only retrain reference
    & --- & retain training data & \tblph & \tblph & \tblph & ---
    & \tblph & \tblph \\
\midrule
Output-channel parents (GA, GradDiff, NPO, SimNPO, IdkDPO, GRU)
    & uniform retain / IDK & \tblph & \tblph & \tblph & \tblph & ---
    & \tblph & \tblph \\
Representation-channel parents (RMU, RepNoise, Circuit Breakers)
    & uniform retain & \tblph & \tblph & \tblph & \tblph & ---
    & \tblph & \tblph \\
\midrule
Crossed protection arms (per parent)
    & \{none, random, matched, mismatched, exact ceiling\}
    & \tblph & \tblph & \tblph & \tblph & \tblph
    & \tblph & \tblph \\
\bottomrule
\end{tabular}
}
\end{table*}

\subsection{Outcomes, Checkpoint Rules, and Statistical Inference}
\label{app:outcomes}

For candidate $x$ at checkpoint $t$, damage is
$d_t(x)=\ell(x;\theta_t)-\ell(x;\theta_0)$, positive values harmful. For a
pool $A$, mean damage is the arithmetic mean and
$\operatorname{CVaR}_{.95}(d_t;A)$ is the mean of the largest
$\max(1,\lceil .05|A|\rceil)$ values; audit pool sizes and tail counts are
always disclosed. Prediction uses each parent's predeclared
terminal-budget checkpoint as the primary horizon, including criterion
failures; fixed-step and criterion-matched horizons are sensitivities.
Within each request we report Spearman correlation, AUROC for realized
top-$K_p$ damage membership, and
$\operatorname{Overlap@}K_p=
|\operatorname{TopK}_{K_p}(r)\cap\operatorname{TopK}_{K_p}(d)|/K_p$
against the chance rate $K_p/|\mathcal C_{\mathrm{audit}}|$.

Protection is a two-part estimand: unconditional criterion reach, then
retention conditional on reach. Single-stage parents use the first saved
reaching checkpoint; repaired arms enter repair there and use the final
saved repair output that still satisfies the same criterion. Conditional
outcomes always report $n_{\mathrm{reach}}/n_{\mathrm{total}}$.

Metrics are first averaged over unlearning seeds within request.
Interaction endpoints use candidate-level bootstrap
(Eq.~\eqref{eq:interaction}); request-level aggregation uses hierarchical
bootstrap over targets and requests, with leave-one-target-out ranges
whenever the top level is small. Paraphrase, generation, extraction, and
general-utility audits are evaluated at selected checkpoints but never
define stopping.

% ==================================================== B. Probe implementation

\section{Probe Implementation and the Fidelity Gate}
\label{app:probe-details}

\subsection{Exact Gradient-Norm Profile}
\label{app:exact-norm}

For candidate $x$ the exact scorer computes $g_x=\nabla_B\ell(x;\theta_0)$
and records $\lVert g_x\rVert_2$. Candidate gradients are processed
sequentially and discarded after their norm is accumulated, so no
$|\mathcal C|\times|B|$ gradient matrix is materialized; the cost is one
candidate-side forward and backward evaluation per example. Exact norm
serves as the output-channel reference predictor, the numerical target for
randomized sensitivity, and the allocation ceiling in the crossed design.

\subsection{Randomized Sensitivity Implementation}
\label{app:randomized-impl}

The randomized scorer draws $R$ seeded Gaussian vectors and normalizes
each over block $B$: $v_r=z_r/\lVert z_r\rVert_2$,
$z_r\sim\mathcal N(0,I_B)$. Directions are generated lazily from the seed
and never materialized as a bank. For every direction, block parameters
are restored from the same saved copy before applying
$\theta_0\pm\eta v_r$; candidate order, batching, padding, masks, and
dropout state are identical across the paired sweeps. The implemented
score is Eq.~\eqref{eq:randomized-susceptibility}; the common
block-dimension factor and square root are omitted as rank-invariant.
Loss differences are accumulated in fp32 even when weights and
activations use lower precision. The profile uses no forget-gradient
backward, no candidate-side backward, and no per-example gradient
materialization; its cost is $2R$ complete candidate-pool forward sweeps.

\subsection{Fidelity Gate: Perturbation Report and
            \texorpdfstring{$A/B/C$}{A/B/C} Decomposition}
\label{app:fidelity-gate}

The gate certifies the randomized profile in two stages, both run before
any confirmatory outcome is unsealed.

\emph{Perturbation report.} For each radius and precision, we measure the
realized perturbation norm $\lVert\eta v\rVert_2/\eta$ and the fraction of
block parameters actually changed by $\pm\eta v$. Low-precision underflow
-- the failure mode that silently zeroes small perturbation components --
appears here as a changed-parameter fraction below one, and disqualifies
the setting before any correlation is computed.

\emph{$A/B/C$ decomposition.} On the diagnostic subset we compare three
quantities per candidate: the exact squared norm
$A=\lVert g_x\rVert_2^2$; the exact projected proxy
$B=(d_B/R)\sum_r (g_x^{\top}v_r)^2$, which shares the probe's Monte Carlo
directions but uses exact inner products; and the finite-difference
estimate $C=r_{\mathrm{rand}}(x)\,d_B$. Rank agreement of $B$ with $A$
isolates Monte Carlo error (relative variance $2/R$, independent of
block dimension); rank agreement of $C$ with $B$ isolates
finite-difference error, including catastrophic cancellation at small
$\eta$. Because the central difference cancels the leading curvature
term, disagreement between $C$ and $B$ at small radii is cancellation
noise, not curvature. The operating $(\eta,R)$ is chosen as the smallest
radius whose $C$--$B$ and $C$--$A$ rank agreements exceed the frozen
threshold with a full perturbation report, then frozen.
Table~\ref{tab:estimator-validation} reports both sides of the gate
boundary.

\subsection{Representation-Proximity Implementation}
\label{app:proximity-impl}

Candidates and forget examples are embedded with mean-pooled final hidden
states of $\theta_0$ over answer tokens; each candidate's score is its
mean cosine similarity to its $k$ nearest forget examples. External
sentence-embedding proximity substitutes a frozen off-the-shelf encoder,
and lexical overlap uses answer-token Jaccard similarity; both are
reported as weaker family members. All three are forward-only and require
no parameter perturbation.

\subsection{Rejected Diagnostic: Forget-Direction Alignment}
\label{app:alignment-impl}

The alignment diagnostic computes one request-level forget gradient in
block $B$, normalizes it, and evaluates candidates at
$\theta_0\pm\eta\hat g_f$ (Eq.~\eqref{eq:directional-probe}). Exact JVP,
vectorized reverse-mode gradient dot, and streaming candidate backward
compute the same local target and are numerical implementations, not
separate hypotheses; their role is to separate finite-difference error
from the scientific failure of the single-direction hypothesis, analyzed
in Appendix~\ref{app:alignment-postmortem}.

% ================================================= C. Protection implementation

\section{Protection Implementation}
\label{app:protection-details}

\subsection{Profile Construction and Pools}
\label{app:pools}

Every scorer covers the declared candidate universe, but only discovery
candidates are eligible for optimization. At fixed budget $K_p$, the
protect pool contains the highest-scoring eligible candidates with
deterministic tie-breaking. Randomized-sensitivity scores are nonnegative:
a high score means high estimated local sensitivity, not positive signed
interference. A size-matched low-response stream $\mathcal R_0$ is drawn
from a frozen near-zero quantile of the discovery fold as the repair-time
reference; for either probe family, a low score does not by itself
establish semantic remoteness. Pool sizes, quantiles, group exclusions,
tie breaking, and fallback rules are fixed without target outcomes, and no
audit candidate can enter either pool.

\subsection{Engine-Then-Repair Pipeline}
\label{app:repair}

Each crossed-design arm runs its parent unmodified from $\theta_0$ to the
first saved checkpoint satisfying the common forget criterion; a parent
that never reaches remains a reach failure and repair is not applied. At
the reaching state, per-sequence and per-answer-token forget losses are
cached as paired references. Repair minimizes answer-token NLL on the
protect pool. At declared refresh steps it computes mean forget and
reference-stream gradients $b_1,\ldots,b_m$ and projects the repair
gradient $v$ away from their span:
\[
    \zeta=(G+\varepsilon I)^{-1}[\langle b_i,v\rangle]_i,
    \qquad G_{ij}=\langle b_i,b_j\rangle,
    \qquad
    \widetilde v=v-\textstyle\sum_i\zeta_i b_i,
\]
with the ridge scaled to the Gram diagonal and a fallback to the primary
forget basis if the declared condition-number threshold is exceeded.
Projection is a local first-order constraint, not a global preservation
guarantee.

\subsection{Paired One-Sided Guard and Auxiliary Bound}
\label{app:guard-proof}

Sequence- and token-level guards (Eq.~\eqref{eq:guard}) have separate
budgets and multipliers. At an audited refresh, a step exceeding either
budget is rejected: the last accepted block parameters and optimizer state
are restored, the repair step size is reduced, and updated multipliers are
retained. Projection is proactive and mean-level; the guard is reactive
and identity-specific. Both require backward evaluations, so the complete
pipeline is not backward-free.

The guard admits a simple tail bound. Let
$a_i=[\bar u_i-\epsilon-u_i]_+$ and suppose an accepted audited checkpoint
satisfies $n^{-1}\sum_i a_i^2\leq\delta^2$. If
$u_i\leq\bar u_i-\epsilon-\Delta$ then $a_i\geq\Delta$, so with
$I_\Delta$ the set of such indices,
\[
    |I_\Delta|\,\Delta^2
    \leq
    \sum_{i\in I_\Delta}a_i^2
    \leq
    n\delta^2
    \qquad\Longrightarrow\qquad
    \frac{|I_\Delta|}{n}\leq\frac{\delta^2}{\Delta^2}:
\]
the fraction of forget examples whose loss regresses by more than
$\Delta$ beyond tolerance is controlled by the accepted guard budget. The
bound applies only to audited loss vectors at accepted refreshes; it is
not a guarantee about generated continuations, extraction, relearning, or
global deletion.

% ============================================ D. Supplementary prediction

\section{Supplementary Prediction Results}
\label{app:prediction-supp}

\subsection{Headline-Probe Secondary Metrics}
\label{app:headline-secondary}

Table~\ref{tab:headline-secondary} reports the secondary prediction
metrics for the two headline probes on every objective, complementing the
$\rho$ matrix of Table~\ref{tab:channel-matrix}. The channel structure
should reappear in each metric: the matched probe dominates AUROC and
Overlap@$K_p$ on its own channel, and tail-restricted $\rho$ verifies that
ranking quality persists inside the damage tail where protection budgets
are spent.

\begin{table*}[t]
\caption{Secondary metrics for the headline probes (TOFU tofu-a180,
sealed audit fold, $n{=}300$, $K_p{=}30$; chance Overlap@$K_p$ is
$0.10$). AUROC and Overlap mirror the channel flip of
Table~\ref{tab:channel-matrix}: the gradient probe is strong on
output-channel objectives (AUROC $0.75$--$0.82$) and at chance on
representation-channel objectives ($0.51$--$0.58$). Tail $\rho$ is
computed within the worst-5\% damage tail ($n{\approx}15$) and is
high-variance; it is reported for completeness and read qualitatively.
$^{\dag}$/$^{\ddag}$ as in Table~\ref{tab:channel-matrix}.}
\label{tab:headline-secondary}
\centering
\scriptsize
\setlength{\tabcolsep}{3pt}
\begin{tabular}{@{}llcccc@{}}
\toprule
Probe & Objective (channel) & $\rho$ & AUROC & Overlap@$K_p$ &
Tail $\rho$ \\
\midrule
Randomized sensitivity & GradDiff (output) & 0.41 & 0.75 & 0.20 & -0.06 \\
& NPO (output) & 0.40 & 0.79 & 0.27 & 0.10 \\
& SimNPO (output) & 0.40 & 0.79 & 0.27 & 0.18 \\
& GRU (output) & 0.43 & 0.77 & 0.23 & -0.10 \\
& GA$^{\ddag}$ (output) & 0.20 & 0.51 & 0.07 & -0.18 \\
& IdkDPO$^{\dag}$ (output) & 0.41 & 0.82 & 0.30 & -0.26 \\
& RMU (repr.) & -0.05 & 0.58 & 0.10 & 0.02 \\
& RepNoise (repr.) & -0.17 & 0.55 & 0.17 & 0.26 \\
& CB$^{\dag}$ (repr.) & -0.25 & 0.53 & 0.17 & -0.12 \\
\addlinespace
Hidden-state proximity & GradDiff (output) & 0.49 & 0.70 & 0.27 & -0.15 \\
& NPO (output) & 0.23 & 0.67 & 0.23 & -0.38 \\
& SimNPO (output) & 0.21 & 0.67 & 0.23 & -0.31 \\
& GRU (output) & 0.42 & 0.68 & 0.23 & -0.18 \\
& GA$^{\ddag}$ (output) & 0.50 & 0.77 & 0.30 & 0.53 \\
& IdkDPO$^{\dag}$ (output) & 0.05 & 0.57 & 0.10 & -0.00 \\
& RMU (repr.) & 0.34 & 0.66 & 0.30 & -0.67 \\
& RepNoise (repr.) & 0.14 & 0.42 & 0.10 & -0.02 \\
& CB$^{\dag}$ (repr.) & 0.09 & 0.66 & 0.17 & -0.03 \\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Randomized-Estimator Accuracy and Stability}
\label{app:estimator-stability}

Table~\ref{tab:estimator-validation} varies one profiling factor at a
time around the frozen operating point; direction-count rows report the
spread across independent direction seeds. The gate rows make the
boundary of Appendix~\ref{app:fidelity-gate} concrete: agreement collapses
in the small-radius cancellation zone and under bf16 underflow, and the
perturbation report flags both before any damage outcome is consulted.
System cost is excluded here (Table~\ref{tab:full-cost}).

\begin{table*}[t]
\caption{Accuracy and stability of the randomized estimator. ``Norm
agreement'' compares randomized sensitivity with exact gradient norm on
rank; ``damage $\rho$'' uses the same sealed outcomes as
Table~\ref{tab:channel-matrix}. Cells report estimate [range over seeds].}
\label{tab:estimator-validation}
\centering
\scriptsize
\setlength{\tabcolsep}{3pt}
\begin{tabular}{@{}llcccc@{}}
\toprule
Factor & Variant & Norm agreement $\rho$ $\uparrow$ &
Norm overlap@$K_p$ $\uparrow$ & Damage $\rho$ $\uparrow$ &
Finite rate $\uparrow$ \\
\midrule
Random directions
    & $R=4$  & \tblph & \tblph & \tblph & \tblph \\
    & $R=16$ & \tblph & \tblph & \tblph & \tblph \\
    & $R=64$ (frozen) & \tblph & \tblph & \tblph & \tblph \\
\midrule
Radius
    & $\eta_0/10$ (cancellation) & \tblph & \tblph & \tblph & \tblph \\
    & $\eta_0$ (frozen) & \tblph & \tblph & \tblph & \tblph \\
    & $10\eta_0$ (curvature) & \tblph & \tblph & \tblph & \tblph \\
\midrule
Precision
    & fp32 weights, fp32 loss & \tblph & \tblph & \tblph & \tblph \\
    & bf16 weights, fp32 loss (underflow) & \tblph & \tblph & \tblph & \tblph \\
\midrule
Parameter block
    & late MLP (frozen) & \tblph & \tblph & \tblph & \tblph \\
    & last layer & \tblph & \tblph & \tblph & \tblph \\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Alignment-Probe Post-Mortem}
\label{app:alignment-postmortem}

The forget-direction alignment probe is numerically correct and
scientifically wrong, and the two facts are established separately.
Table~\ref{tab:local-fidelity} shows that symmetric FD recovers the exact
JVP directional derivative to high precision, so its failure in
Table~\ref{tab:channel-matrix} is not estimator error.
Table~\ref{tab:score-anatomy} decomposes the local response on the
diagnostic gradient subset: gradient magnitude carries the transferable
signal on output-channel damage, whereas the signed-alignment factor --
the part specific to the initial forget direction -- does not survive the
trajectory. Table~\ref{tab:trajectory-mechanism} completes the account:
saved-checkpoint displacement rotates away from the initial direction,
and alignment's predictive value decays with that rotation.

\begin{table*}[t]
\caption{Numerical fidelity of implementations targeting the initial
forget-direction derivative. Exact JVP is the reference; this table
establishes local correctness only. Relative error excludes reference
values inside the frozen zero tolerance.}
\label{tab:local-fidelity}
\centering
\small
\setlength{\tabcolsep}{4pt}
\begin{tabular}{@{}lcccccc@{}}
\toprule
Estimator & Candidate reverse & Abs.\ error $\downarrow$ &
Rel.\ error $\downarrow$ & Sign agreement $\uparrow$ &
JVP rank $\rho$ $\uparrow$ & JVP overlap@$K_p$ $\uparrow$ \\
\midrule
Exact JVP
    & no & --- & --- & 1.000 & 1.000 & 1.000 \\
Symmetric FD
    & no & \tblph & \tblph & \tblph & \tblph & \tblph \\
One-sided FD
    & no & \tblph & \tblph & \tblph & \tblph & \tblph \\
Vmap gradient dot
    & yes & \tblph & \tblph & \tblph & \tblph & \tblph \\
Streaming backward
    & yes & \tblph & \tblph & \tblph & \tblph & \tblph \\
\bottomrule
\end{tabular}
\end{table*}

\begin{table*}[t]
\caption{Score anatomy on the frozen gradient subset. Cells report
within-request Spearman correlation with terminal damage, one
representative objective per channel. Magnitude transfers on the output
channel; the alignment factor does not; neither carries the
representation channel.}
\label{tab:score-anatomy}
\centering
\scriptsize
\setlength{\tabcolsep}{3.2pt}
\begin{tabular}{@{}lccccc@{}}
\toprule
Signal & Formula & Candidate backward &
GradDiff (output) & GRU (output) & RMU (repr.) \\
\midrule
Initial NLL
    & $\ell(x;\theta_0)$ & no & \tblph & \tblph & \tblph \\
Answer length
    & $|x_{\mathrm{answer}}|$ & no & \tblph & \tblph & \tblph \\
\midrule
Gradient magnitude
    & $\|g_x\|_2$ & yes & \tblph & \tblph & \tblph \\
Signed alignment
    & $\cos(g_x,g_f)$ & yes & \tblph & \tblph & \tblph \\
Magnitude $\times$ alignment
    & $\langle g_x,\hat g_f\rangle$ & yes & \tblph & \tblph & \tblph \\
Directional FD
    & $\delta_\eta(x;\hat g_f)$ & no & \tblph & \tblph & \tblph \\
Absolute directional FD
    & $|\delta_\eta(x;\hat g_f)|$ & no & \tblph & \tblph & \tblph \\
\bottomrule
\end{tabular}
\end{table*}

\begin{table*}[t]
\caption{Checkpoint-level directional persistence. $a_t$ is the signed
cosine between saved block displacement and the initial forget direction.
``Alignment pred.''\ is directional-FD correlation with damage at that
checkpoint; ``norm pred.''\ is exact-norm correlation with the same
damage. A positive relationship between $a_t$ and alignment prediction
supports, but does not prove, the rotation explanation.}
\label{tab:trajectory-mechanism}
\centering
\scriptsize
\setlength{\tabcolsep}{3pt}
\begin{tabular}{@{}lccccc@{}}
\toprule
Parent & Units & $a_t$ [IQR] &
$\rho(a_t,\text{alignment pred.})$ &
Norm pred.\ [IQR] & Alignment pred.\ [IQR] \\
\midrule
GradDiff & \tblph & \tblph & \tblph & \tblph & \tblph \\
NPO+retain & \tblph & \tblph & \tblph & \tblph & \tblph \\
RMU & \tblph & \tblph & \tblph & \tblph & \tblph \\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Specificity Controls}
\label{app:specificity}

Randomized sensitivity does not use the target forget request, so its
unlearning-specific reading requires controls. We compare every frozen
profile against damage from target unlearning, a different forget
request, benign supervised fine-tuning, and random motion matched to the
target displacement norm within the same block.

\begin{table*}[t]
\caption{Specificity controls. Cells report candidate-level correlation
with realized damage. Similar prediction across all columns supports the
narrower interpretation of general update sensitivity rather than
request-specific unlearning risk.}
\label{tab:specificity}
\centering
\small
\setlength{\tabcolsep}{4pt}
\begin{tabular}{@{}lcccc@{}}
\toprule
Predictor & Target unlearning & Other forget request &
Benign SFT & Matched random update \\
\midrule
Exact gradient norm & \tblph & \tblph & \tblph & \tblph \\
Randomized sensitivity & \tblph & \tblph & \tblph & \tblph \\
Hidden-state proximity & \tblph & \tblph & \tblph & \tblph \\
Forget-direction alignment & \tblph & \tblph & \tblph & \tblph \\
Random ranking & \tblph & \tblph & \tblph & \tblph \\
\bottomrule
\end{tabular}
\end{table*}

% ============================================ E. Supplementary protection

\section{Supplementary Protection Results}
\label{app:protection-supp}

\subsection{Protection-Budget Sweep}
\label{app:budget-sweep}

The default-budget crossed result is in Table~\ref{tab:crossed}.
Table~\ref{tab:budget-sweep} varies the eligible pool fraction inside the
same output-channel parent (GradDiff entry state) with processed repair
tokens fixed, so a larger pool does not receive a larger optimization
budget. The channel prediction is that the matched selectors' advantage
persists across budgets while random allocation dilutes.

\begin{table*}[t]
\caption{Matched protection-budget sweep (GradDiff parent). Every selector
uses the same entry checkpoint, repair algorithm, updates, processed
tokens, and sealed native audit. Reach is unconditional; remaining
outcomes use the final saved criterion-satisfying repair output.}
\label{tab:budget-sweep}
\centering
\scriptsize
\setlength{\tabcolsep}{3pt}
\begin{tabular}{@{}lccccc@{}}
\toprule
Selector & $K_p/|\mathcal C_{\mathrm{disc}}|$ & Reach $\uparrow$ &
Native mean $\Delta$NLL $\downarrow$ &
Native CVaR$_{.95}$ $\downarrow$ & Utility ret.\ $\uparrow$ \\
\midrule
Exact gradient norm & .05 & \tblph & \tblph & \tblph & \tblph \\
Exact gradient norm & .10 & \tblph & \tblph & \tblph & \tblph \\
Exact gradient norm & .20 & \tblph & \tblph & \tblph & \tblph \\
\midrule
Randomized sensitivity & .05 & \tblph & \tblph & \tblph & \tblph \\
Randomized sensitivity & .10 & \tblph & \tblph & \tblph & \tblph \\
Randomized sensitivity & .20 & \tblph & \tblph & \tblph & \tblph \\
\midrule
Hidden-state proximity & .05 & \tblph & \tblph & \tblph & \tblph \\
Hidden-state proximity & .10 & \tblph & \tblph & \tblph & \tblph \\
Hidden-state proximity & .20 & \tblph & \tblph & \tblph & \tblph \\
\midrule
Random & .05 & \tblph & \tblph & \tblph & \tblph \\
Random & .10 & \tblph & \tblph & \tblph & \tblph \\
Random & .20 & \tblph & \tblph & \tblph & \tblph \\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Repair-Component Ablations}
\label{app:repair-ablations}

Table~\ref{tab:protection-ablations} ablates projection and the paired
guard within each parent's matched-selector arm. The prediction is that
selection quality (the crossed contrast) and repair safety (the guard) are
separable: removing the guard should trade forget regression for
retention, and removing projection should slow repair without changing
which candidates benefit.

\begin{table*}[t]
\caption{Repair-component ablations at the frozen budget, matched selector
per parent. Conditional outcomes report
$n_{\mathrm{reach}}/n_{\mathrm{total}}$. Paired drift is measured only at
audited refreshes.}
\label{tab:protection-ablations}
\centering
\scriptsize
\setlength{\tabcolsep}{2.3pt}
\begin{tabular}{@{}llcccccccc@{}}
\toprule
Parent & Repair arm & Proj. & Guard & Reach $\uparrow$ &
Forget recall $\downarrow$ & Native mean $\downarrow$ &
Native CVaR $\downarrow$ & Utility $\uparrow$ &
Paired drift $\downarrow$ \\
\midrule
GradDiff & No repair
    & --- & --- & \tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
GradDiff & Full repair
    & on & on & \tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
GradDiff & No projection
    & off & on & \tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
GradDiff & No paired guard
    & on & off & \tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
\midrule
RMU & No repair
    & --- & --- & \tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
RMU & Full repair
    & on & on & \tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
RMU & No paired guard
    & on & off & \tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Strong-Method Context}
\label{app:strong-methods}

Table~\ref{tab:main} places the complete channel-matched pipeline
alongside strong recent unlearning methods under the same evaluation
contract. It is context rather than the confirmatory claim: the pipeline
is a layer over a parent objective, so the causal comparison remains the
crossed design of Table~\ref{tab:crossed}, and no channel-blind method can
exhibit the interaction that constitutes our endpoint. Initial and
retain-only retraining checkpoints are behavioral references, not
budget-matched competitors.

\begin{table*}[t]
\caption{Strong-method comparison on TOFU single-author requests. Reach
includes all requests. Single-stage methods use their first saved reaching
checkpoint; repaired methods use their final saved output that still
satisfies the same criterion. Conditional outcomes report
$n_{\mathrm{reach}}/n_{\mathrm{total}}$; lower native-audit damage is
better.}
\label{tab:main}
\centering
\scriptsize
\setlength{\tabcolsep}{3.0pt}
\begin{tabular}{@{}lcccccc@{}}
\toprule
Method & Reach $\uparrow$ & Forget recall $\downarrow$ &
Para.\ recall $\downarrow$ & Native mean $\Delta$NLL $\downarrow$ &
Native CVaR$_{.95}$ $\downarrow$ & Utility ret.\ $\uparrow$ \\
\midrule
Initial target reference &
\tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
Retain-only retrain reference &
\tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
\midrule
GA & \tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
GradDiff & \tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
NPO+retain & \tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
IdkDPO & \tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
SimNPO & \tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
GRU & \tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
RMU & \tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
RepNoise & \tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
Circuit Breakers & \tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
GUARD & \tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
Cheng split-aware &
\tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
\midrule
Channel-matched probe + guarded repair (ours) &
\tblph & \tblph & \tblph & \tblph & \tblph & \tblph \\
\bottomrule
\end{tabular}
\end{table*}

% ============================================== F. Cost, replication, stress

\section{Cost, Replication, and Stress Details}
\label{app:full-transfer}

\subsection{Profiling-Cost Protocol}
\label{app:cost-protocol}

All profiling comparisons use identical candidate order, token batches,
hardware, warm-up, model precision, and loss accumulation. Timing starts
before method-specific shared setup and ends after all scores are
persisted; direction-specific methods include their forget-gradient
evaluation, which randomized sensitivity does not require. The systems
claim is limited to eliminating candidate-side reverse mode and
per-example gradient materialization; runtime speedup is measured, not
assumed.

\begin{table*}[t]
\caption{Complete-pool profiling cost (Qwen2.5-7B, bf16, batch 4). $R$ is
the random-direction count; ``pool evaluations'' reports the actual
traversal form rather than equating one candidate backward with one
full-pool forward sweep. Cells report median [IQR].}
\label{tab:full-cost}
\centering
\scriptsize
\setlength{\tabcolsep}{2.6pt}
\begin{tabular}{@{}lccccccc@{}}
\toprule
Implementation & $R$ & Pool evaluations & Forget reverse &
Candidate reverse & Wall time & Peak memory & Tokens/s \\
\midrule
Exact gradient norm
    & --- & $|\mathcal C|\times(\mathrm{F{+}B})$ & no & $|\mathcal C|$
    & \tblph & \tblph & \tblph \\
Randomized sensitivity
    & 16 & $32\times$forward & no & 0
    & \tblph & \tblph & \tblph \\
Randomized sensitivity
    & 64 & $128\times$forward & no & 0
    & \tblph & \tblph & \tblph \\
Forget-direction FD
    & --- & $2\times$forward & 1 & 0
    & \tblph & \tblph & \tblph \\
Exact JVP
    & --- & $1\times$forward-mode JVP & 1 & 0
    & \tblph & \tblph & \tblph \\
Vmap gradient dot
    & --- & chunked F+B & 1 & yes
    & \tblph & \tblph & \tblph \\
Streaming backward
    & --- & $|\mathcal C|\times(\mathrm{F{+}B})$ & 1 & $|\mathcal C|$
    & \tblph & \tblph & \tblph \\
Hidden-state proximity
    & --- & representation forward & no & 0
    & \tblph & \tblph & \tblph \\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Replication and Stress Protocol}
\label{app:replication}

Each replication dataset (WMDP-bio, MUSE-News, RWKU) reruns the reduced
crossed roster with identical endpoints, channel assignments, probe
settings, and fold discipline; only the dataset adapter changes. Because
each dataset shifts the probes' base rates -- forget--retain lexical and
embedding proximity distributions are disclosed per dataset -- the
confirmatory quantity is the interaction, not any absolute correlation.
Stress settings follow the same protocol where the honesty clauses
predict attenuation: PISTOL's maximal structural entanglement inflates
proximity scores on candidates that outputs-channel updates do not damage,
and MUSE-Books tests whether verbatim-memorization removal behaves as an
output-channel member as its loss definition declares.

\begin{table*}[t]
\caption{Replication and stress results. Each row reports the
preregistered interaction $\Delta$ (Eq.~\eqref{eq:interaction}) with
candidate-bootstrap CI, the crossed-protection outcome (matched CVaR
lowest per parent), and the dataset's proximity profile disclosed for
base-rate context.}
\label{tab:full-transfer-results}
\centering
\scriptsize
\setlength{\tabcolsep}{3pt}
\begin{tabular}{@{}lccccc@{}}
\toprule
Setting & Units & Forget--retain proximity profile &
Prediction $\Delta$ [95\% CI] & Matched wins (GradDiff) &
Matched wins (RMU) \\
\midrule
TOFU (primary) & 20 requests & \tblph & \tblph & \tblph & \tblph \\
WMDP-bio & 1 corpus & \tblph & \tblph & \tblph & \tblph \\
MUSE-News & 5 requests & \tblph & \tblph & \tblph & \tblph \\
RWKU & 10 requests & \tblph & \tblph & \tblph & \tblph \\
MUSE-Books (stress) & 3 requests & \tblph & \tblph & \tblph & \tblph \\
PISTOL (stress) & 10 requests & \tblph & \tblph & \tblph & \tblph \\
\bottomrule
\end{tabular}
\end{table*}

% ====================================================== G. Additional limits

\section{Additional Limitations and Reporting Scope}
\label{app:limitations}

Randomized sensitivity measures local update sensitivity and may predict
benign fine-tuning or unrelated updates as well as target unlearning;
Table~\ref{tab:specificity} determines how far the interpretation can be
narrowed. The score depends on the declared block, radius, perturbation
distribution, precision, loss, and checkpoint -- changing any of these
defines another profile, certified separately by the fidelity gate. On
hardware with efficient vectorized per-example gradients, wall-clock
savings over exact profiling are not guaranteed; the claim is
backward-freedom and memory, reported in Table~\ref{tab:full-cost}.
Neither the profile nor guarded repair certifies deletion: removal claims
require direct, paraphrase, extraction, relearning, and privacy audits,
and profile-defined pools are never the sole evidence of successful
unlearning or preservation.
